---
title: "Getting Started with RBiblioSynth"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with RBiblioSynth}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

demo_biblio <- utils::head(RBiblioSynth::biblio_sample, 120)
demo_source_file <- tempfile(fileext = ".csv")
utils::write.csv(demo_biblio, demo_source_file, row.names = FALSE)
demo_sources <- list(
  sample = list(
    file = demo_source_file,
    db = "generic",
    format = "csv"
  )
)
```

# Getting Started with RBiblioSynth

This vignette provides a comprehensive introduction to RBiblioSynth, a modular R framework for bibliometric analysis and systematic review reporting.

## Installation

```r
# Load the package from a local development checkout
# devtools::load_all()
library(RBiblioSynth)
```

## Quick Start

### 1. Load Data

RBiblioSynth supports multiple bibliographic databases:

```r
# Or use the sample dataset
demo_biblio <- RBiblioSynth::biblio_sample
```

### 2. Run the official pipeline

```{r, eval = FALSE}
# Run the public M0-M3 pipeline
pipeline_result <- run_pipeline(
  demo_sources,
  modules = c("m0", "m1", "m2", "m3"),
  config = biblio_config(
    counting_mode = "full",
    dedup_method = "title_year_normalized",
    report_format = "none"
  ),
  export = FALSE
)

m0_result <- pipeline_result$modules$m0
bib_data <- m0_get_bib_data(m0_result)

# Access organized data from M0
countries <- m0_get(m0_result, "countries")
authors <- m0_get(m0_result, "authors")
annual <- m0_get(m0_result, "annual")
```

### 3. Run M1 Main Information

```r
m1_result <- pipeline_result$modules$m1

# View results
print(m1_result$data$overview)
print(m1_result$data$authors)
print(m1_result$data$bradford)

# Access plots
m1_result$artifacts$plots$overview$m1_overview_bar
m1_result$artifacts$plots$lotka$m1_lotka_scatter
```

### 4. Run M2 Annual Production

```r
# Module result returned by the pipeline
m2_result <- pipeline_result$modules$m2

# View growth model comparison
print(m2_result$data$growth_models$comparison)

# View best model
print(m2_result$data$growth_models$best_model)

# Access forecast
print(m2_result$data$forecasting$arima)
```

### 5. Run M3 Countries

```r
# Module result returned by the pipeline
m3_result <- pipeline_result$modules$m3

# View spatial statistics
print(m3_result$data$spatial$morans_i)

# View economic correlation
print(m3_result$data$economic)

# View collaboration indices
print(m3_result$data$collaboration_indices)
```

## Module-Level Entry Points

If you want to run modules individually, the public API remains available:

```{r, eval = FALSE}
m0_result <- run_m0(demo_sources, export = FALSE)
bib_data <- m0_get_bib_data(m0_result)
annual <- m0_get(m0_result, "annual")

m1_result <- run_m1(bib_data, export = FALSE)
m2_result <- run_m2(annual, export = FALSE)
m3_result <- run_m3(bib_data, export = FALSE)
```

## Advanced Features

### Bootstrap Confidence Intervals

All key metrics include bootstrap confidence intervals:

```r
# Compute h-index with bootstrap CI
citations <- c(10, 25, 5, 15, 30, 8, 12, 45, 3, 18)

h_index <- function(x) {
  sorted <- sort(x, decreasing = TRUE)
  max(which(sorted >= seq_along(sorted)))
}

ci <- bootstrap_ci(citations, h_index, R = 1000, method = "bca")
print(ci)
```

### Reproducible report bundle

```{r, eval = FALSE}
cfg <- biblio_config(report_format = "latex_bundle")
pipeline_result <- run_pipeline(demo_sources, config = cfg)

# Generated bundle paths
pipeline_result$artifacts$files
```

### Spatial Statistics

```r
# Moran's I for spatial autocorrelation
m3_result$data$spatial$morans_i

# LISA clusters
m3_result$data$spatial$lisa$clusters

# Getis-Ord Gi* hot spots
m3_result$data$spatial$getis_ord$hotspots
```

### QAP Network Test

```r
# Test correlation between collaboration and citation networks
collab_matrix <- m3_result$data$collaboration_indices$matrix
citation_matrix <- compute_citation_correlation_matrix(bib_data)

qap_result <- m3_compute_qap_test(
  collab_matrix,
  citation_matrix,
  n_permutations = 1000
)

print(qap_result$observed_correlation)  # 0.45
print(qap_result$p_value_two_tailed)    # 0.002
print(qap_result$interpretation)
# "The networks show a moderate positive correlation (r = 0.450), 
# which is significant (p = 0.0020)."
```

## PRISMA Flow Diagrams

```r
# Define PRISMA specification
prisma <- list(
  title = "Systematic Review Title",
  identification = list(
    records_database = 1250,
    records_other = 45,
    duplicates_removed = 180
  ),
  screening = list(
    records_screened = 1115,
    excluded_screening = 780
  ),
  eligibility = list(
    fulltext_assessed = 335,
    excluded_fulltext = 195
  ),
  included = list(
    studies_included = 140
  )
)

# Generate PRISMA diagram
m0_render_prisma_diagram(prisma)
```

## Export

```r
# Export to JSON
write_json_artifact(m1_result$data$overview, "output/m1_overview.json")

# Export plots
export_plot_artifact(m1_result$artifacts$plots$overview$m1_overview_bar, 
                     "output/overview", width =10, height = 8)
```

## Statistical Tests

Each module includes formal hypothesis testing:

```r
# M1 Hypothesis tests
m1_result$data$hypotheses

# M2 Hypothesis tests
m2_result$data$hypotheses

# M3 Hypothesis tests  
m3_result$data$hypotheses
```

## Next Steps

1. Explore the [Module Reference](../reference/index.html) for detailed function documentation
2. Use `run_pipeline()` for the official end-to-end workflow
3. Use module-level entrypoints when you need custom orchestration

## Citation

If you use RBiblioSynth in your research, please cite:

```bibtex
@article{rbibliosynth2026,
  title = {RBiblioSynth: A Comprehensive R Framework for Bibliometric Analysis},
  author = {Mayorga},
  journal = {SoftwareX},
  year = {2026}
}
```
